### 2015-12-28 Projects/Exploratory_Convective_PPE/Report/convective_ppe_heating.tex
convective report edits

To accomplish this, NIST will examine documented on-duty injuries and fatalities of firefighters due to thermal exposure of PPE used by structural firefighters.  Further NIST will conduct laboratory thermal testing of commercially available PPE as part of this phase of the study:
\begin{enumerate}
\item NIST shall examine documented on-duty injuries and fatalities of firefighters due to thermal exposure of structural firefighting PPE \cite{Weinschenk:Chicago,Overholt:San_Francisco,Weinschenk:PGCounty}.
\item NIST shall conduct laboratory thermal testing of commercially available structural firefighting PPE as part of this phase of the study, with a focus on convective heat transfer.  The testing shall include any relevant performance requirements for structural firefighting PPE thermal exposure in NFPA 1971.
\item NIST shall conduct full-scale thermal testing of structural firefighting PPE at field test sites.

\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
There is a gap in the understanding of the thermal failure temperature (melting point) of various pieces of personal protective equipment (PPE) worn or used by firefighters.  As a result, there may be a misunderstanding of the thermal conditions in which a firefighter can safely operate.  Ignition and melt testing on various pieces of PPE will define the thermal failure temperatures and lead to a better understanding of the conditions of a safer firefighting environment.  Although there are flame and heat resistance tests in NFPA 1971, Standard on Protective Ensembles for Structural Fire Fighting and Proximity Fire Fighting and NFPA 1981, Standard on Open-Circuit Self Contained Breathing Apparatus for Fire and Emergency Services, these tests do not provide the engineering data needed to connect laboratory based research data with field data (thermally damaged equipment) from actual fire incidents.

@@

The initial phase of this study would examine documented on-duty injuries and fatalities of firefighters due to thermal exposure of structural firefighting PPE.  This phase of the study would also work with the National Fire Protection Association (NFPA) 1971 Standard on Protective Ensembles for Structural Fire Fighting technical committee on ways to enhance the operational effectiveness of structural firefighting PPE.  Initial laboratory thermal testing of commercially available structural firefighting PPE shall also be conducted as part of this phase of the study.

Existing Tests – high thermal flux, short exposure
TTP Test- PPE
Thermo man – (pyro man) PPE
Flame contact – helmet
While under bench scale test conditions it is less difficult to decouple convective heating from a radiant heat exposure.  However with current test apparatus while convective heating is dominant, there is some level of radiant heating that is contributing to the heat transfer.
Under full scale conditions in a flow path the PPE is exposed to a combination of convective and radiative heating.
LoDDs
Use - Cherry Road and San Francisco as examples
Items examined

@@

Future Research

Flow loop    with 20 mph and 300 to 500 F - Donnelly
Substrates
2D vs 3D samples
Gaps – refer to Lawson and Stroup
Instrumentation – ref to Vettori
Steam

\chapter{Discussion}

>>>>

To accomplish this, NIST will examine documented on-duty injuries and fatalities of firefighters due to thermal exposure of PPE used by structural firefighters. Further NIST will conduct laboratory thermal testing of commercially available PPE as part of this phase of the study:
\begin{enumerate}
\item NIST shall examine documented on-duty injuries and fatalities of firefighters due to thermal exposure of structural firefighting PPE \cite{Weinschenk:Chicago,Overholt:San_Francisco,Weinschenk:PGCounty}.
\item NIST shall conduct laboratory thermal testing of commercially available structural firefighting PPE as part of this phase of the study, with a focus on convective heat transfer. The testing shall include any relevant performance requirements for structural firefighting PPE thermal exposure in NFPA 1971.
\item NIST shall conduct full-scale thermal testing of structural firefighting PPE at field test sites.
\end{enumerate}

\chapter{Introduction}
\pagenumbering{arabic}
\setcounter{page}{1}
There is a gap in the understanding of the thermal failure temperature (melting point) of various pieces of personal protective equipment (PPE) worn or used by firefighters. As a result, there may be a misunderstanding of the thermal conditions in which a firefighter can safely operate. Ignition and melt testing on various pieces of PPE will define the thermal failure temperatures and lead to a better understanding of the conditions of a safer firefighting environment. Although there are flame and heat resistance tests in NFPA 1971, Standard on Protective Ensembles for Structural Fire Fighting and Proximity Fire Fighting~\cite{nfpa1971_2013} and NFPA 1981, Standard on Open-Circuit Self Contained Breathing Apparatus for Fire and Emergency Services~\cite{nfpa1981_2013}, these tests do not provide the engineering data needed to connect laboratory based research data with field data (thermally damaged equipment) from actual fire incidents.

@@

The initial phase of this study examines the documented on-duty injuries and fatalities of firefighters due to thermal exposure of structural firefighting PPE. This phase of the study would also work with the National Fire Protection Association (NFPA) 1971 Standard on Protective Ensembles for Structural Fire Fighting technical committee on ways to enhance the operational effectiveness of structural firefighting PPE. Initial laboratory thermal testing of commercially available structural firefighting PPE shall also be conducted as part of this phase of the study.

\chapter{Prior On-Duty Injuries and Fatalities}

There have been many previous fire incidents~\cite{NIOSH:Pettit,NIOSH:Washenitz,NIOSH:Mezzanotte,NIOSH:McFall,NIOSH:McFall2,NIOSH:McFall3,NIOSH:Berardinelli,NIOSH:Koedam,NIOSH:McFall4,NIOSH:Tarley,NIOSH:Braddee,NIOSH:Merinar,NIOSH:Bowyer2,NIOSH:Loflin,NIOSH:Bowyer} in which changes in the flow paths are thought to have had an adverse impact on firefighter and occupant safety. Table~\ref{tab:LODD} lists the NIOSH investigation reports from the past 15 years in which it could be determined that a flow path played a role in the related incident. This table lists the NIOSH report number, the outcome, and a brief description of the flow path details.

Based on a review of these incidents, it is clear that fires with rapidly developing or changing ventilation may lead to flow paths that are a significant hazard to the fire service during a response. The development of (or changes to) a flow path could be caused by the failure of a component of the structure, such as a door, window, or portion of a ceiling, wall or floor. Environmental conditions such as wind can generate hazardous thermal conditions within a flow path. Uncoordinated ventilation procedures can also be the cause of increased thermal hazards within a flow path.

\begin{table}[!ht]
\caption[Flow path related LODD/LODI incidents.]
{Flow path related LODD/LODI incidents.}
\begin{tabular}{lll}
\toprule
NIOSH Report No. & No. of LODDs/LODIs & Flow Path Details                                              \\
\midrule
99-F01   \cite{NIOSH:Pettit}        &  3 LODDs            &  From apartment into hallway on 10th       \\
                                    &                     &  floor of high-rise apartment building     \\
99-F21   \cite{NIOSH:Washenitz}     &  2 LODDs            &  Basement to 1st floor                     \\
                                    &  2 LODIs            &                                            \\
F2000-04 \cite{NIOSH:Mezzanotte}    &  3 LODDs            &  1st floor to 2nd floor                    \\
                                    &  3 civilian deaths  &                                            \\
F2000-16 \cite{NIOSH:McFall}        &  1 LODD             &  2nd floor hallway through                 \\
                                    &  1 LODI             &  2nd floor apartment                       \\
                                    &  1 civilian death   &                                            \\
F2000-23 \cite{NIOSH:McFall2}       &  1 LODD             &  From ground level to 1st floor then to    \\
                                    &  2 LODIs            &  2nd floor, flow exited through ceiling    \\
F2000-43 \cite{NIOSH:McFall3}       &  1 serious LODI     &  1st floor to 2nd floor                    \\
                                    &  2 other LODIs      &                                            \\
F2004-02 \cite{NIOSH:Berardinelli}  &  1 LODD             &  1st floor to basement                     \\
F2005-02 \cite{NIOSH:Koedam}        &  1 LODD             &  Rear to front of the building             \\
                                    &  4 LODIs            &                                            \\
F2005-04 \cite{NIOSH:McFall4}       &  1 LODD             &  Basement to 1st floor                     \\
                                    &  9 LODIs            &                                            \\
F2007-09 \cite{NIOSH:Tarley}        &  1 LODD             &  3 story training burn - flow through      \\
                                    &  2 LODIs            &  all levels                                \\
F2007-35 \cite{NIOSH:Braddee}       &  4 LODIs            &  1st floor to 2nd floor                    \\
F2009-11 \cite{NIOSH:Merinar}       &  2 LODDs            &  Rear to front of the building             \\
F2011-13 \cite{NIOSH:Bowyer2}       &  2 LODDs            &  Lower level up stairs and through         \\
                                    &                     &  entry door and garage                     \\
F2011-31 \cite{NIOSH:Loflin}        &  1 LODD             &  Fire extended from lower level apartment  \\
F2012-28 \cite{NIOSH:Bowyer}        &  1 LODD             &  Attic fire extended into closed           \\
                                    &  1 LODI             &  porch and then into 2nd floor             \\
\bottomrule
\end{tabular}
\label{tab:LODD}
\end{table}

\chapter{Existing Tests}
Existing Tests - high thermal flux, short exposure

\section{Thermal Protective Performance}
TTP Test - PPE

\section{ThemoMan}
Thermo man - (pyro man) PPE

\section{Flame Contact}
Flame contact - helmet

While under bench scale test conditions it is less difficult to decouple convective heating from a radiant heat exposure.  However with current test apparatus while convective heating is dominant, there is some level of radiant heating that is contributing to the heat transfer.
Under full scale conditions in a flow path the PPE is exposed to a combination of convective and radiative heating.
LoDDs
Use - Cherry Road and San Francisco as examples
Items examined

@@

\chapter{Future Research}
Flow loop  with 20 mph and 300 to 500 F - Donnelly
Substrates
2D vs 3D samples
Gaps - refer to Lawson and Stroup
Instrumentation - ref to Vettori
Steam
